---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
dat <- read.csv("data/activity.csv")
dat$date <- as.Date(as.character(dat$date), "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

### 1. Total number of steps per day
```{r daily.steps}
daily.steps <- aggregate(steps ~ date, data=dat, FUN=sum, na.rm=T)

```
### 2. Histogram of the total number of steps taken each day

```{r hist}
hist(daily.steps$steps, col="blue", xlab="Steps per day", ylab="Step count", main="")
```

### 3. Summary statistics of daily steps

```{r summary}
summary(daily.steps$steps)
```

## What is the average daily activity pattern?

### 1. 5 minute interval plot against step count across all days

``` {r time.series}

interval.data <- aggregate(steps ~ interval, data=dat, FUN=mean, na.rm=T)
with(interval.data, plot(steps~interval, type="l", col="blue", xlab="Interval Number", ylab="Step count averaged over days", main="Averaged step counts for every time interval"))
```

### 2. 5-minute interval with maximum average step count.

``` {r max.interval}

max.interval <- interval.data[which(interval.data$steps == max(interval.data$steps)),]

```

The max interval is interval number `r max.interval$interval` having a maximum average step count value of `r round(max.interval$steps, 2)`.

## Imputing missing values

### 1. Total number of missing values in the datase

``` {r na}
na.count <- sum(is.na(dat))
```

There are `r na.count` rows with missing values.

## Are there differences in activity patterns between weekdays and weekends?
